Memory-aware Adaptive Scheduling of Scientific Workflows On Heterogeneous Architectures\\

### 2024-10-31 main-disk-model.tex

    Hence, during the task prioritization phase in~\cite{sulaiman2021hybrid},  the standard deviation of the computation cost
    (between processors) is computed, and added to the mean value to account for the differences between processor speeds.
    In the processor choice phase, the entry task and the longest parent tasks are duplicated 
    during idle times on the processor.

@@

    Rahman~\etal~\cite{rahman2013}'s dynamic critical path (DCP) algorithm for grids maps tasks to machines
    by calculating the critical path in the graph dynamically at every step.
    %For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
    %on the start time of a task (differing by the slack this task has).
    %All tasks on this critical path have the same earliest and latest start times, because they cannot be delayed.
    They schedule the first task on the critical path to the best suitable processor and recompute the critical path.
    %The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
    %If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum
    %execution time for the task is chosen.
    The heuristic also uses the same processor to schedule parent and children tasks, as to avoid data transfer between processors.
    The approach is evaluated on random workflows of the size up to 300 tasks.

@@

\skug{
\section{Model} %: \skug{Full: 4, polished: 3}}
\label{sec:model}

CHAGES IN THIS CHAPTER
}
    We first describe the model for our target applications, which are (large scientific) workflows for which we do not have perfect a priori knowledge,
    in Section~\ref{sec.mod.work}.  Next, we define the execution
    environment, a heterogeneous system (in terms of processor speed and memory size),
    in Section~\ref{sec.mod.plat}. Finally, we present the optimization problem in
    Section~\ref{sec.mod.pb}. The key notation is summarized in Table~\ref{tabnotation}.

    \subsection{Workflow}
    \label{sec.mod.work}
    A workflow is modeled as a directed acyclic graph $G=(V, E)$, where $V$ is the set of vertices (tasks), and
    $E$ is a set of directed edges of the form $e=(u,v)$, with $u,v\in V$, expressing precedence constraints between tasks.
    Each task~$u \in V$  is performing $w_u$ operations, and it also
    requires some amount of memory to be executed, denoted as~$m_u$.
    Each edge $e=(u,v) \in E$ has a cost~$c_{u,v}$ that corresponds to the size of the output file written by task~$u$ and used as input by task~$v$.

    Note that $m_u$ is the total memory usage
    of a task during its execution, including input and output files currently being read and written,
    and hence the total memory requirement  for the execution of task~$u$ consists of the maximum
    between the input files
    (total size of the files to be received from the parents),
    the output files (total size of the files to be sent to the children),
    and the total memory size~$m_u$ (usually achieving the maximum):
    \[
        r_u = \max\left\{m_u , \sum_{v:(v,u)\in E}c_{v,u}, \sum_{v:(u,v)\in E} c_{u,v}\right\}.
    \]

    The parents of a task~$u\in V$ are the directly preceding tasks that must be completed before $u$ can be started, i.e., the set of parents is
    $ \parents{u} = \{v \in V: (v,u) \in E\}$. A task without parents is called a {\it source task}.
    The children tasks of~$u$ are the tasks following~$u$ directly according to the precedence constraints, i.e.,
    $ \children{u} = \{v \in V: (u,v) \in E\}$. A task without children is called a {\it target task}.
    Each task may have multiple parents and children.

Furthermore, we place ourselves in a context where we do not have perfect knowledge
of the task parameters ($w_u$ and $m_u$) before the tasks start their execution,
but only estimates~\cite{rahman2013,GARG2015256}.  
%. \AB{Add motivation: related work with variable task durations for instance...}
Hence, scheduling decisions are made on these estimated parameters, and
may be reconsidered at runtime when a task starts its execution and we know its exact parameters.
\AB{Variability only on $w_u$ for now}

    \subsection{Execution environment}
    \label{sec.mod.plat}
\skug{changes here}
The goal is to execute the workflow on a heterogeneous system, denoted as $\cluster$, which
consists of $k$ processors $p_1, \dots, p_k$.
Each processor $p_u$ ($1 \leq u \leq k$) has an individual memory of size $M_u$ and a speed~$s_u$.
All processors have access to a shared disk of unlimited size, where they can write
and read data: Processor~$p_u$ has a bandwidth $\bw_u$ (resp. $\br_u$) to write  (resp. read),
and hence it takes a time $\frac{c_{u,v}}{bw_u}$ to write the data produced for~$p_v$, 
if this data is evicted from memory.

%, but of much slower speed.
All communications happen over the disk - after one processor has finished writing data there, another one can read
it into its individual memory.
We can decide to evict some data from the individual memory onto the disk to free space in the individual memory.
This data can later be read back into memory of the same or any other processpr at any time.
\skug {check this statement: If data has been written on disk, it is removed from the local memory (no two copies can be kept at the same time).}

If the entire memory required by the task ~$u\in V$ fits in the free memory of a processor~$p_j$,
    then the execution time of this task on this processor is expressed as $\frac{w_u}{s_j}$.
If the memory requirement of the task exceeds the available memory on the processor, the task can still be executed there,
    albeit slower.
The part of memory requirement that exceeds the available memory will be offloaded to the disk, slowing down the execution.

    \[
        w_{real} = \frac{w_u}{s_j} +    \frac{q_s \times w_u}{s_j} \frac{m_i - M_j}{m_i}
    \]
where $q_s$ sit he slowdown coefficient and $(m_i - M_j)/m_i$ is the part of the task memory that needs to be offloaded.
    \skug{check this equation}
%We assume that all processors are connected with the same bandwidth~$\beta$.
%    \hmey{Maybe mention that variable bandwidths are part of future work...?}

Multiple processors can communicate with the disk in parallel, but the communication of each processor  with the disk is sequential.
The processor can read and write at the same time, but only one file can be read and only one written at any point in time.
We keep track of the current ready times of each processor $j$, ready time for computation $\rt_j^c$ (when one task has finished
computing and another one can start),
ready time for writing $\rt_j^w$ (when a file has been written to disk and another file can be written), and ready time for
 reading+$\rt_j^r$.
Initially, all the ready times are set to~$0$.
We also keep track of the currently available memory, $availM_j$ on the processor memory.

Furthermore, $\PD_j$ is a priority queue with the {\em pending data}
that are in the memory of size $\MM_j$ but may be evicted to the disk
more memory is needed on~$p_j$.
They are ordered by non-decreasing size and correspond to some $c_{u,v}$.
When scheduling a tass, the decision has to be made to execute this task on a processor with enough memory
    without slowdown, but with possibly more time to read the input files, or to execute the same task
    on a processor with less memory (and hence a slowdown), but potentially with a higher processor speed or with less times
    spent reading files.

We use the \algo{memDag} algorithm developed by Kayaaslan \etal~\cite{KAYAASLAN20181} to compute
the memory requirement; it transforms the workflow into a series-parallel graph
and then finds the traversal that leads to the minimum memory consumption.

\skug{Todo: describe deviations}

    \begin{table}
        \begin{center}
            \begin{tabular}{rl}
                \hline
                \textbf{Symbol}                       & \textbf{Meaning}                                         \\
                \hline
                $G = (V, E)$                          & Workflow graph, set of vertices (tasks) and edges        \\
                $\parents{u}$, $\children{u}$         & Parents of a task $u$, children of a task $u$            \\
                $m_u$                                 & Memory weight of task $u$                                \\
                $w_u$                                 & Workload of task $u$  (normalized execution time)          \\
                $c_{u,v}$                             & Communication volume along the edge $(u,v)\in E$         \\
                $F$, $\mathcal{F}$                    & A partitioning function and the partition it creates     \\
                $V_i$                                 & Block number $i$                                         \\ %\wrt~some $F$   \\
                $\cluster$, $k$                    & Computing system and its number of processors           \\
                $p_j$, proc($V_i$)                          & Processor number $j$, processor of block $V_i$                 \\
                $M_j$, $MC_j$, $s_j$                               & Memory size, comm. buffer size, and speed of proc.\ $p_j$                          \\
                $\beta$                     & Bandwidth in the compute system                                \\
                $\bottomlevel{u}$                      & Bottom weight of task $u$ \\
                $\mu_G$, $\mu_i$ & Makespan of the entire workflow $G$ and of a block $V_i$               \\
                $\Gamma = (\mathcal{V}, \mathcal{E})$                      & Quotient graph, its vertices and its edges        \\
                $r_u$, $r_{V_i}$                            & Memory requirement of task $u$ and of block $V_i$                 \\
                \hline
            \end{tabular}
        \end{center}
        \caption{Notation} \label{tabnotation}
    \end{table}

>>>>

    Hence, during the task prioritization phase in~\cite{sulaiman2021hybrid},  the standard deviation of the computation cost
    (between processors) is computed, and added to the mean value to account for the differences between processor speeds.
    In the processor choice phase, the entry task and the longest predecessor tasks are duplicated 
    during idle times on the processor.

@@

    Rahman~\etal~\cite{rahman2013}'s dynamic critical path (DCP) algorithm for grids maps tasks to machines
    by calculating the critical path in the graph dynamically at every step.
    %For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
    %on the start time of a task (differing by the slack this task has).
    %All tasks on this critical path have the same earliest and latest start times, because they cannot be delayed.
    They schedule the first task on the critical path to the best suitable processor and recompute the critical path.
    %The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
    %If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum
    %execution time for the task is chosen.
    The heuristic also uses the same processor to schedule predecessor and successor tasks, as to avoid data transfer between processors.
    The approach is evaluated on random workflows of the size up to 300 tasks.

@@

\section{Model} %: \skug{Full: 4, polished: 3}}
\label{sec:model}

%\skug{
%CHANGES IN THIS CHAPTER
%}
    We first describe the model for our target applications, which are (large scientific) workflows for which we do not have perfect a priori knowledge,
    in Section~\ref{sec.mod.work}.  Next, we define the execution
    environment, a heterogeneous system (in terms of processor speed and memory size),
    in Section~\ref{sec.mod.plat}. Finally, we present the optimization problem in
    Section~\ref{sec.mod.pb}. The key notation is summarized in Table~\ref{tabnotation}.

    \subsection{Workflow}
    \label{sec.mod.work}
    A workflow is modeled as a directed acyclic graph $G=(V, E)$, where $V$ is the set of vertices (tasks), and
    $E$ is a set of directed edges of the form $e=(u,v)$, with \mbox{$u,v\in V$}, expressing precedence constraints between tasks.
    Each task~$u \in V$  is performing $w_u$ operations, and it also
    requires some amount of memory to be executed, denoted as~$m_u$.
    Each edge $e=(u,v) \in E$ has a cost~$c_{u,v}$ that corresponds to the size of the output file written by task~$u$ and used as input by task~$v$.

    Note that $m_u$ is the total memory usage
    of a task during its execution, including input and output files currently being read and written,
    and hence it is greater than the total size of input files
    (received from the predecessor tasks) and than 
    the total size of  output files (sent to successor tasks): 
    $$   m_u \geq \max \left\{ \sum_{v:(v,u)\in E}c_{v,u}, \sum_{v:(u,v)\in E} c_{u,v} \right\} . $$

%     the total memory requirement  for the execution of task~$u$ consists of the maximum
%    between the input files
%    (total size of the files to be received from the parents),
%    the output files (total size of the files to be sent to the children),
%    and the total memory size~$m_u$ (usually achieving the maximum):
%    \[
%        r_u = \max\left\{m_u , \sum_{v:(v,u)\in E}c_{v,u}, \sum_{v:(u,v)\in E} c_{u,v}\right\}.
%    \]

    The predecessors of a task~$u\in V$ are the directly preceding tasks that must be completed before $u$ can be started, i.e., the set of predecessors is
    $ \parents{u} = \{v \in V: (v,u) \in E\}$. A task without predecessors is called a {\it source task}.
    The successors of a task~$u$ are the tasks following~$u$ directly according to the precedence constraints, i.e.,
    $ \children{u} = \{v \in V: (u,v) \in E\}$. A task without successors is called a {\it target task}.
    Each task may have multiple predecessors and successors.

Furthermore, we place ourselves in a context where we do not have perfect knowledge
of the length of the tasks, i.e., $w_u$, % ($w_u$ and $m_u$) 
before the tasks complete their execution,
but only estimates~\cite{rahman2013,GARG2015256}.  
%. \AB{Add motivation: related work with variable task durations for instance...}
Hence, scheduling decisions are made on these estimated parameters, and
may be reconsidered at runtime when a task completes its execution. 
%when a task starts its execution and we know its exact parameters.
%\AB{Variability only on $w_u$ for now}

    \subsection{Execution environment}
    \label{sec.mod.plat}

%\skug{changes here}
The goal is to execute the workflow on a heterogeneous system, denoted as $\cluster$, which
consists of $k$ processors $p_1, \dots, p_k$.
Each processor $p_j$ ($1 \leq j \leq k$) has an individual memory of size $M_j$ and a speed~$s_j$.
All processors have access to a shared disk of unlimited size, where they can write
and read data: Processor~$p_j$ has a bandwidth $\bw_j$ (resp.~$\br_j$) to write  (resp. read),
and hence it takes a time $\frac{c_{u,v}}{bw_j}$ to write the data produced by task~$u$
on~$p_j$ for task~$v$, 
if this data is evicted from memory.

%, but of much slower speed.
All communications happen over the disk -- after one processor has finished writing data there, another one can read
it to load the data on its individual memory. 
Hence, we can decide to evict some data from the individual processor memories if it has been 
written onto the disk, 
in order to free some space in the individual memories.
This data can later be read back into memory of the same or any other processor at any time.
%\skug {check this statement: If data has been written on disk, it is removed from the local memory (no two copies can be kept at the same time).} \AB{We can keep two copies: evict if we need to }

If the entire memory required by  task ~$u\in V$ fits in the available memory of processor~$p_j$,
    then the execution time of this task on this processor is expressed as $\frac{w_u}{s_j}$.
However, if the memory requirement of the task exceeds the available memory on the processor, 
then task can still be executed there, but it will be slowed down. Indeed, 
%    albeit slower.
the part of memory requirement that exceeds the available memory on processor~$j$, $availM_j$, 
will be offloaded to the disk, slowing down the execution:

    \[
        w_{real} = \frac{w_u}{s_j} +  q_s \times   (m_u - availM_j) , %   \frac{q_s \times w_u}{s_j} \frac{m_i - M_j}{m_i}
    \]
where $q_s$ is the slowdown coefficient, and $(m_u - availM_j)$ is the part 
of the task memory that needs to be offloaded.
\AB{Svetlana please check, I think we said in the end that we use absolute value 
of the amount of memory not fitting locally...}
%    \skug{check this equation}
%We assume that all processors are connected with the same bandwidth~$\beta$.
%    \hmey{Maybe mention that variable bandwidths are part of future work...?}

Multiple processors can communicate with the disk in parallel, but the communication of each processor  with the disk is sequential.
The processor can read and write at the same time, but only one file can be read and only one written at any point in time.
We keep track of the current ready times of each processor $j$, ready time for computation $\rt_j^c$ (when one task has finished
computing and another one can start),
ready time for writing $\rt_j^w$ (when a file has been written to disk and another file can be written), and ready time for
 reading~$\rt_j^r$.
Initially, all the ready times are set to~$0$.
We also keep track of the currently available memory, $availM_j$ on the processor memory.

Furthermore, $\PD_j$ is a priority queue with the {\em pending data}
that are in the memory of processor~$p_j$,  but that may be evicted to the disk
if more memory is needed on~$p_j$.
They are ordered by non-decreasing size and correspond to some files~$c_{u,v}$.
When scheduling a task, the decision has to be made to execute this task on a processor with enough memory
without slowdown, but with possibly more time to read the input files, or to execute the same task
on a processor with less memory (and hence a slowdown), but potentially with a higher processor speed 
or with less times spent reading files.

We use the \algo{memDag} algorithm developed by Kayaaslan \etal~\cite{KAYAASLAN20181} to compute
the memory requirement; it transforms the workflow into a series-parallel graph
and then finds the traversal that leads to the minimum memory consumption.
\AB{When do we need this? Probably more in Algorithms, right? }

%\skug{Todo: describe deviations}

    \begin{table}
        \begin{center}
            \begin{tabular}{rl}
                \hline
                \textbf{Symbol}                       & \textbf{Meaning}                                         \\
                \hline
                $G = (V, E)$                          & Workflow graph, set of vertices (tasks) and edges        \\
                $\parents{u}$, $\children{u}$         & Predecessors of a task $u$, successors of a task $u$            \\
                $m_u$                                 & Memory weight of task $u$                                \\
                $w_u$                                 & Workload of task $u$  (normalized execution time)          \\
                $c_{u,v}$                             & Communication volume along the edge $(u,v)\in E$         \\
                $F$, $\mathcal{F}$                    & A partitioning function and the partition it creates     \\
                $V_i$                                 & Block number $i$                                         \\ %\wrt~some $F$   \\
                $\cluster$, $k$                    & Computing system and its number of processors           \\
                $p_j$, proc($V_i$)                          & Processor number $j$, processor of block $V_i$                 \\
                $M_j$, $MC_j$, $s_j$                               & Memory size, comm. buffer size, and speed of proc.\ $p_j$                          \\
                $\beta$                     & Bandwidth in the compute system                                \\
                $\bottomlevel{u}$                      & Bottom weight of task $u$ \\
                $\mu_G$, $\mu_i$ & Makespan of the entire workflow $G$ and of a block $V_i$               \\
                $\Gamma = (\mathcal{V}, \mathcal{E})$                      & Quotient graph, its vertices and its edges        \\
                $r_u$, $r_{V_i}$                            & Memory requirement of task $u$ and of block $V_i$                 \\
                \hline
            \end{tabular}
        \end{center}
        \caption{Notation} \label{tabnotation}
    \end{table}
